---
title: "Homework 3"
author: "Riya Maliekal"
format: html
editor: visual
---

# The Effect of Recommendation on Choice
- loading the data 
- We use logistic regression to explain the data as it has a binary output of whether a recommendation is followed or not (1,0) as the output variable. 


```{r}
data=read.csv("C:/Users/ASUS/Documents/GitHub/Homework_assignment/recommendations.csv")
```

## Fit the logistic regression model. 
```{r}
#|eval: FALSE
suppressWarnings(library(tidyverse))
model=glm(RecommendationFollowed ~ Mode, data = data, family=binomial)
summary(model)
```


## Create a new data frame to use for plotting
```{r}
new_data <- data.frame(Mode = c("Auditory", "Visual"))
```

## Predict the probabilities of following the recommendation for each mode

```{r}
new_data$prob <- predict(model, new_data, type = "response")
```

## Create the plot
```{r}
ggplot(new_data, aes(x = Mode, y = prob)) +
  geom_bar(stat = "identity", fill= 'purple', width=0.25) +
  labs(x = "Recommendation Mode", y = "Probability of Following Recommendation", title = "Effect of Recommendation Mode on Whether a Recommendation is Followed") +
  theme_minimal()
```


